---
title: "Model Estimation and Model Diagnostics"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This tutorial is created using R [markdown](https://rmarkdown.rstudio.com/) and [knitr](https://yihui.name/knitr/). It illustrates how to use the GDINA R pacakge (version `r packageVersion("GDINA")`) for various CDM analyses.

## Model Estimation

The following code estimates the G-DINA model. For extracting item and person parameters from G-DINA model, please see [this tutorial](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/GDINA_example.html).
```{r}
library(GDINA)
dat <- sim10GDINA$simdat
Q <- matrix(c(1,0,0,
              0,1,0,
              0,0,1,
              1,0,1,
              0,1,1,
              1,1,0,
              1,0,1,
              1,1,0,
              1,1,1,
              1,0,1),byrow = T,ncol = 3)

est <- GDINA(dat = dat, Q = Q, model = "GDINA", verbose = 0)
```



## Q-matrix validation

The **Qval()** function is used for Q-matrix validation. By default, it implements de la Torre and Chiu's (2016) algorithm. The following example use the stepwise method (Ma & de la Torre, 2019) instead.
```{r}
Qv <- Qval(est, method = "Wald")
Qv
```
To further examine the q-vectors that are suggested to be modified, you can draw the mesa plots (de la Torre & Ma, 2016):

```{r}
plot(Qv, item = 9)
plot(Qv, item = 10)
```

We can also examine whether the G-DINA model with the suggested Q had better relative fit:
```{r}
sugQ <- extract(Qv, what = "sug.Q")
est.sugQ <- GDINA(dat, sugQ, verbose = 0)
anova(est,est.sugQ)
```

## Item-level model comparison
Based on the suggested Q-matrix, we perform item level model comparison using the Wald test (see de la Torre, 2011; de la Torre & Lee, 2013; Ma, Iaconangelo & de la Torre, 2016) to check whether any reduced CDMs can be used. Note that score test and likelihood ratio test (Sorrel, Abad, Olea, de la Torre, and Barrada, 2017; Sorrel, de la Torre, Abad, & Olea, 2017; Ma & de la Torre, 2018) may also be used.
```{r}
mc <- modelcomp(est.sugQ)
mc
```

We can fit the models suggested by the Wald test based on the rule in Ma, Iaconangelo and de la Torre (2016) and compare the combinations of CDMs with the G-DINA model:
```{r}
est.wald <- GDINA(dat, sugQ, model = extract(mc,"selected.model")$models, verbose = 0)
anova(est.sugQ,est.wald)
```

## Absolute fit evaluation
The test level absolute fit include M2 statistic, RMSEA and SRMSR (Maydeu-Olivares, 3013; Liu, Tian, & Xin, 2016; Hansen, Cai, Monroe, & Li, 2016; Ma, 2019) and the item level absolute fit include log odds and transformed correlation (Chen, de la Torre, & Zhang, 2013), as well as heat plot for item pairs.
```{r}
# test level absolute fit
mft <- modelfit(est.wald)
mft
# item level absolute fit
ift <- itemfit(est.wald)
ift
summary(ift)
plot(ift)
```


The estimated latent class size can be obtained by
```{r}
extract(est.wald,"posterior.prob")
```


The tetrachoric correlation between attributes can be calculated by
```{r}
# psych package needs to be installed
library(psych)
psych::tetrachoric(x = extract(est.wald,"attributepattern"),
                   weight = extract(est.wald,"posterior.prob"))
```

## Classification Accuracy
The following code calculates the test-, pattern- and attribute-level classification accuracy indices based on GDINA estimates using approaches in Iaconangelo (2017) and Wang, Song, Chen, Meng, and Ding (2015).
```{r}
CA(est.wald)
```
## References

Chen, J., de la Torre, J., & Zhang, Z. (2013). Relative and Absolute Fit Evaluation in Cognitive Diagnosis Modeling.
_Journal of Educational Measurement, 50_, 123-140.

de la Torre, J., & Lee, Y. S. (2013). Evaluating the wald test for item-level comparison of saturated and reduced models in cognitive diagnosis. *Journal of Educational Measurement, 50*, 355-373.

de la Torre, J., & Ma, W. (2016, August). Cognitive diagnosis modeling: A general framework approach and its implementation in R. A short course at the
fourth conference on the statistical methods in Psychometrics, Columbia University, New York.

Hansen, M., Cai, L.,  Monroe, S., & Li, Z. (2016). Limited-information goodness-of-fit testing of diagnostic classification item response models. *British Journal of Mathematical and Statistical Psychology. 69,* 225--252.

Iaconangelo, C.(2017). *Uses of Classification Error Probabilities in the Three-Step Approach to Estimating Cognitive Diagnosis Models.* (Unpublished doctoral dissertation). New Brunswick, NJ: Rutgers University.

Liu, Y., Tian, W., & Xin, T. (2016). An Application of M2 Statistic to Evaluate the Fit of Cognitive Diagnostic Models. *Journal of Educational and Behavioral Statistics, 41*, 3-26.

Ma, W. (2019). Evaluating the fit of sequential G-DINA model using limited-information measures. *Applied Psychological Measurement*.

Ma, W. & de la Torre, J. (2018). Category-level model selection for the sequential G-DINA model. *Journal of Educational and Behavorial Statistics*.

Ma,W., & de la Torre, J. (2019). An empirical Q-matrix validation method for the sequential G-DINA model. *British Journal of  Mathematical and Statistical Psychology*. 

Ma, W., Iaconangelo, C., & de la Torre, J. (2016). Model similarity, model selection and attribute classification. *Applied Psychological Measurement, 40*, 200-217.

Maydeu-Olivares, A. (2013). Goodness-of-Fit Assessment of Item Response Theory Models. *Measurement, 11*, 71-101.

Sorrel, M. A., Abad, F. J., Olea, J., de la Torre, J., & Barrada, J. R. (2017). Inferential Item-Fit Evaluation in Cognitive Diagnosis Modeling. *Applied Psychological Measurement, 41,* 614-631.

Sorrel, M. A., de la Torre, J., Abad, F. J., & Olea, J. (2017). Two-Step Likelihood Ratio Test for Item-Level Model Comparison in Cognitive Diagnosis Models. *Methodology, 13*, 39-47.

Wang, W., Song, L., Chen, P., Meng, Y., & Ding, S. (2015). Attribute-Level and Pattern-Level Classification Consistency and Accuracy Indices for Cognitive Diagnostic Assessment.
*Journal of Educational Measurement, 52* , 457-476.

```{r}
sessionInfo()
```
