---
title: "Higher-order GDINA Estimation"
author: "Wenchao"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

This tutorial is created using R [markdown](https://rmarkdown.rstudio.com/) and [knitr](https://yihui.name/knitr/). It illustrates how to use the GDINA R pacakge (version `r packageVersion("GDINA")`) to estimate the higher-order G-DINA model.

### Model Estimation

The following code estimates the higher-order G-DINA model.
```{r}
library(GDINA)

# A simulated data in GDINA package
dat <- sim10GDINA$simdat
Q <- sim10GDINA$simQ

# Estimating GDINA model
est <- GDINA(dat = dat, Q = Q, model = "GDINA", att.dist = "higher.order", higher.order = list(model = "2PL"))
```

### Summary Information

The following code extracts the summary information from higher-order GDINA estimates.
```{r}
#####################################
#
#      Summary Information
# 
#####################################
# print estimation information
est

# summary information
summary(est)

AIC(est) #AIC
BIC(est) #BIC
logLik(est) #log-likelihood value
deviance(est) # deviance: -2 log-likelihood
npar(est) # number of parameters
nobs(est) # number of observations
```
You can use *extract* with argument *discrim* to extract discrimination indices. The first column gives $P(1)-P(0)$ and the second column gives the GDINA discrimination index.
```{r}
# discrimination indices
extract(est, "discrim")
```

### Model Parameters

```{r}
#####################################
#
#      structural parameters
# 
#####################################
```
The following code gives the item probalities of each reduced latent classes. As shown below, the probability of answering item 1 correctly for individuals who do not master the required attribute is `r coef(est)[[1]][1]`, and the probability of answering item 1 correctly for individuals who master the required attribute is `r coef(est)[[1]][2]`:
```{r}
coef(est) # item probabilities of success for each reduced latent class
```
The following code gives the item probalities of each reduced latent classes with standard errors.
```{r}
coef(est, withSE = TRUE) # item probabilities of success & standard errors

```
The following code gives delta parameters.
```{r}
coef(est, what = "delta") # delta parameters
```
The following code gives delta parameters with standard errors. 
```{r}
coef(est, what = "delta", withSE = TRUE) # delta parameters

```
The following code gives $P(0)$ and 1-$P(0)$, which is guessing and slipping parameters.
```{r}
coef(est, what = "gs") # guessing and slip parameters
```
The following code gives guessing and slipping parameters with standard errors. 
```{r}
coef(est, what = "gs", withSE = TRUE) # guessing and slip parameters & standard errors

# Estimated proportions of latent classes
```
The following code gives the slope and intercept parameters for attributes. As you can see, the estimated slope and intercept for the first attribute are `r coef(est,"lambda")[1,1]` and `r coef(est,"lambda")[1,2]`, respectively:
```{r}
coef(est,"lambda")
# success probabilities for each latent class
```
The following code gives item success probabilities for all latent classes,
```{r}
coef(est,"LCprob")

#####################################
#
#      person parameters
# 
#####################################
```
The following code returns EAP estimates of attribute patterns (for the first six individuals). As you can see, the EAP estimate of attribute profile for the first individual is (`r personparm(est)[1,]`):
```{r}
head(personparm(est)) # EAP estimates of attribute profiles
```
 By specifying *what* argument, the following code gives MAP estimates of attribute patterns (for the first six individuals).
```{r}
head(personparm(est, what = "MAP")) # MAP estimates of attribute profiles
```
 The following code extracts MLE estimates of attribute patterns (for the first six individuals).
```{r}
head(personparm(est, what = "MLE")) # MLE estimates of attribute profiles
```

### Some Plots
```{r}
#####################################
#
#           Plots
# 
#####################################

#plot item response functions for item 10
```
The following code gives item response functions of item 10.
```{r}
plot(est, item = 10)
```

The following code gives item response functions of item 10 with error bars.
```{r}
plot(est, item = 10, withSE = TRUE) # with error bars

```

The following code plots mastery probabilities of three attributes for individuals 1,20 and 50.
```{r}
#plot mastery probability for individuals 1, 20 and 50
plot(est, what = "mp", person = c(1, 20, 50))
```

### Advanced Topics
```{r}
#####################################
#
#      Advanced elements
# 
#####################################

head(indlogLik(est)) # individual log-likelihood
head(indlogPost(est)) # individual log-posterior
extract(est,"designmatrix") #design matrix
extract(est,"linkfunc") #link functions

sessionInfo()
```
