---
title: "Model-data fit"
author: "Wenchao Ma"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

This tutorial is created using R [markdown](https://rmarkdown.rstudio.com/) and [knitr](https://yihui.name/knitr/). It illustrates how to use the GDINA R package (version `r packageVersion("GDINA")`) for model-data fit evaluation.

We use the ECPE data for illustration and fit the GDINA model to the data first.

### Model Estimation

The following code estimates the G-DINA model.
```{r}
library(GDINA)

dat <- realdata_ECPE$dat
Q <- realdata_ECPE$Q

# Estimating GDINA model
est <- GDINA(dat = dat, Q = Q, model = "GDINA", mono.constraint = TRUE)
```

### Model fit at test-level

The model-data fit at test level can be obtained using `modelfit()` function.
This function calculates $M_2$ statistic for G-DINA model with dichotmous responses (Liu, Tian, & Xin, 2016; Hansen, Cai, Monroe, & Li, 2016) and for sequential G-DINA model with graded responses (Ma, 2020). It also calculates SRMSR and RMSEA2.


```{r}
modelfit(est)
```
Interestingly,  `itemfit()` also provide model-data fit at test level. By default,
it prints the maximum z statistics for transformed correlation and log odds ratio
for item-pairs, which quantify the "worst" fit.
```{r}
itemfit(est)
```

## Model-data fit at item-pair level

`itemfit()` also provide model-data fit at item-pair level. The heatmap plots for
transformed correlation and log odds ratio show the p-values of test statistics for
all item pairs. A significant (red) test statistic indicates some misfit.

```{r}
ift <- itemfit(est)
plot(ift)
```

### Model data fit at item level

Model data fit can also be evaluated at item level. The code below gives the maximum
z statistics for transformed correlation and log odds ratio at item level and thus can 
be viewed as the "worst" fit for item pairs related to that specific item.
```{r}
ift <- itemfit(est)
ift$max.itemlevel.fit
```

Item fit can also be evaluated using `itemfitPD()` function. This function calculate item fit statistics from the power-divergence family (Najera, Ma, Sorrel and Abad, 2025).
This function can use bootstrap methods with Stone correction. Parallel computing is supported
on computers with multiple cores.

```{r cache=TRUE}
ipd <- itemfitPD(est, bootstrap = TRUE, Stone = TRUE, R = 1000, cores=10)
ipd
```

To find the test statistics and p-value, the following code can be used:
```{r}
ipd$PD
```

```{r}
sessionInfo()
```


### References

Chen, J., de la Torre, J., & Zhang, Z. (2013). Relative and Absolute Fit Evaluation in Cognitive Diagnosis Modeling. Journal of Educational Measurement, 50, 123-140.

Hansen, M., Cai, L., Monroe, S., & Li, Z. (2016). Limited-information goodness-of-fit testing of diagnostic classification item response models. British Journal of Mathematical and Statistical Psychology. 69, 225–252.

Liu, Y., Tian, W., & Xin, T. (2016). An Application of M2 Statistic to Evaluate the Fit of Cognitive Diagnostic Models. Journal of Educational and Behavioral Statistics, 41, 3-26.

Ma, W. (2020). Evaluating the fit of sequential G-DINA model using limited-information measures. Applied Psychological Measurement, 44, 167-181.

Ma, W., & de la Torre, J. (2020). GDINA: An R Package for Cognitive Diagnosis Modeling. Journal of Statistical Software, 93(14), 1-26.

Maydeu-Olivares, A. (2013). Goodness-of-Fit Assessment of Item Response Theory Models. Measurement, 11, 71-101.

Nájera, P., Ma, W., Sorrel, M.A. et al. Assessing item-level fit for the sequential G-DINA model. Behaviormetrika (2025). https://doi.org/10.1007/s41237-025-00263-8

