---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# GDINA Package for Cognitively Diagnostic Analyses

[![Project Status: Active ? The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![Build Status](https://travis-ci.org/Wenchao-Ma/GDINA.svg?branch=master)](https://travis-ci.org/Wenchao-Ma/GDINA)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/GDINA)](https://cran.r-project.org/package=GDINA)
[![](https://cranlogs.r-pkg.org/badges/GDINA?color=brightgreen)](https://cran.r-project.org/package=GDINA)
[![](http://cranlogs.r-pkg.org/badges/grand-total/GDINA?color=green)](https://cran.r-project.org/package=GDINA)

## How to cite the package

Ma, W. & de la Torre, J. (2019). GDINA: The generalized DINA model framework. R package version 2.4. Retrived from https://CRAN.R-project.org/package=GDINA

Visit the package website https://wenchao-ma.github.io/GDINA for examples and tutorials.

## Learning resources

* [NCME digital module 5](https://ncme.elevate.commpartners.com/) on the G-DINA model and the use of graphical user interface for CDM analyses

* [NCME training session](https://www.ncme.org/meetings/annualmeeting) on April 4, 2019


## Features of the package

*  Estimating G-DINA model and a variety of widely-used models subsumed by the G-DINA model, including the DINA model, DINO model, additive-CDM (A-CDM), linear logistic model (LLM), reduced reparametrized unified model (RRUM), multiple-strategy DINA model for dichotomous responses
*  Estimating models within the G-DINA model framework using user-specified design matrix and link functions
*  Estimating Bugs-DINA, DINO and G-DINA models for dichotomous responses
*  Estimating sequential G-DINA model for ordinal and nominal responses
*  Estimating the generalized multiple-strategy cognitive diagnosis models (experimental)
*  Estimating the diagnostic tree model (experimental)
*  Modelling independent, saturated, higher-order, loglinear smoothed, and structured joint attribute distribution
*  Accommodating multiple-group model analysis
*  Imposing monotonic constrained success probabilities
*  Accommodating binary and polytomous attributes
*  Validating Q-matrix under the general model framework
*  Evaluating absolute and relative item and model fit
*  Comparing models at the test and item levels
*  Detecting differential item functioning using Wald and likelihood ratio test
*  Simulating data based on all aforementioned CDMs
*  Providing graphical user interface for users less familiar with R


## Installation

To install this package from source:

1) Windows users may need to install the
   [Rtools](https://CRAN.R-project.org/bin/windows/Rtools/) and include the checkbox option of installing Rtools to their path for 
   easier command line usage. Mac users will have to
   download the necessary tools from the [Xcode](https://itunes.apple.com/ca/app/xcode/id497799835?mt=12) and its
   related command line tools (found within Xcode's Preference Pane under Downloads/Components); most Linux
   distributions should already have up to date compilers (or if not they can be updated easily). 

2) Install the `devtools` package (if necessary), and install the package from the Github source code.

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("Wenchao-Ma/GDINA")
```

The stable version of GDINA should be installed from R CRAN at [here](https://CRAN.R-project.org/package=GDINA)



