---
title: "A quick reference to GDINA R package"
author: "Wenchao Ma"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A quick reference to GDINA R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[utf8]{inputenc}
---

## What can this package do?

This package provides a framework for conducting a variety of cognitively diagnostic analyses
for dichotomous and polytomous responses.

For more details, see [here](https://wenchao-ma.github.io/GDINA/index.html)


## How do we start to use this package?

* To estimate various CDMs, use **GDINA()** function
* To extract item and person parameters from GDINA estimates, use **coef()** and **personparm()** functions, respectively
* To evaluate absolute model fit, use **modelfit()** or **itemfit()** functions
* To evaluate relative model fit, use **AIC()**, **BIC()** or **deviance()** functions
* To validate Q-matrix, use **Qval()** function
* To detect differential item functioning, use **dif()** function
* To run graphical user interface, use **startGDINA()** function
* To simulate data from various CDMs, use **simGDINA()** function

Visit [here](https://wenchao-ma.github.io/GDINA/reference/index.html) for online reference

## A quick example

Below is an illustration showing some analyses that you could do using the package. You can copy the code, paste and run it in R console.

1. Load the package and assume we have the following item response matrix and Q-matrix:
```{r,eval=FALSE}
library(GDINA)
dat <- sim10GDINA$simdat
Q <- matrix(c(1,0,0,
              0,1,0,
              0,0,1,
              1,0,1,
              0,1,1,
              1,1,0,
              0,0,1,
              1,0,0,
              1,1,1,
              1,0,1),byrow = T,ncol = 3)
```

2. Estimate the G-DINA model:
```{r,eval=FALSE}
est <- GDINA(dat = dat, Q = Q, model = "GDINA")
```


3. Conduct the Q-matrix validation. By default, it implements de la Torre and Chiu's (2016) algorithm using a fixed cutoff. This is fast but sometimes suggests too many modifications.
```{r,eval=FALSE}
Qv <- Qval(est)
Qv
```
To avoid using fixed cutoffs and also take uncertainty in item parameter estimation into account, you may consider the stepwise method:
```{r,eval=FALSE}
Qv2 <- Qval(est,method = "Wald")
Qv2
```

To further examine the q-vectors that are suggested to be modified, you can draw mesa plots:

```{r,eval=FALSE}
plot(Qv, item = 9)
```

4. Perform item-level model selection to see if the saturated G-DINA model can be simplified:
```{r,eval=FALSE}
mc <- modelcomp(est)
mc
```

5. Assess model-data fit at test and item levels:
```{r,eval=FALSE}
# test level absolute fit
mft <- modelfit(est)
mft
# item level absolute fit
ift <- itemfit(est)
ift
summary(ift)
plot(ift)
```

6. Calculate classification accuracy
```{r,eval=FALSE}
CA(est)
```

## More Examples

* [G-DINA model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/GDINA_example.html) 
* [LCDM model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/LCDM_example.html) 
* [Higher-order GDINA model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/HOGDINA_example.html)  

If you would like to contribute an example to this website, please send me your .Rmd file.
