---
title: "A quick reference to GDINA R package"
author: "Wenchao Ma"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A quick reference to GDINA R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[utf8]{inputenc}
---

## What can this package do?

This package provides a framework for conducting a series of cognitive diagnosis analysis for dichotomous and polytomous responses.

For more details, see [here](https://wenchao-ma.github.io/GDINA/index.html)

For an example of how to use the package, see [here](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/CDMAnalysis_example.html)

## How to use this package?

* To estimate various CDMs, use **GDINA()** function
* To extract item and person parameters from GDINA estimates, use **coef()** and **personparm()** functions, respectively
* To extract other estimation components, use **extract()** function
* To evaluate absolute model fit, use **modelfit()**, **itemfit()** or **itemfitPD()** function
* To evaluate relative model fit, use **AIC()**, **BIC()** or **deviance()** function
* To validate Q-matrix, use **Qval()** function
* To detect differential item functioning, use **dif()** function
* To regress estimated attribute or profile classifications on covariates with three-step ML correction, use **ThreeStepCov()**
* To run graphical user interface, use **startGDINA()** function
* To simulate data from various CDMs, use **simGDINA()** function

Visit [here](https://wenchao-ma.github.io/GDINA/reference/index.html) for an online reference

## A quick example

Below is an illustration showing some analyses that you could do using the package. You can copy the code, paste and run it in R console. For a more comprehensive example, see [this tutorial](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/CDMAnalysis_example.html).

1. Load the package and assume we have the following item response matrix and Q-matrix:
```{r,eval=FALSE}
library(GDINA)
dat <- sim10GDINA$simdat
Q <- matrix(c(1,0,0,
              0,1,0,
              0,0,1,
              1,0,1,
              0,1,1,
              1,1,0,
              0,0,1,
              1,0,0,
              1,1,1,
              1,0,1),byrow = T,ncol = 3)
```

2. Estimate the G-DINA model:
```{r,eval=FALSE}
est <- GDINA(dat = dat, Q = Q, model = "GDINA")
```


3. Conduct the Q-matrix validation. By default, it implements de la Torre and Chiu's (2016) algorithm using a fixed cutoff. This is fast but sometimes suggests too many modifications.
```{r,eval=FALSE}
Qv <- Qval(est)
Qv
```
To avoid using fixed cutoffs and also take uncertainty in item parameter estimation into account, you may consider the stepwise method:
```{r,eval=FALSE}
Qv2 <- Qval(est,method = "Wald")
Qv2
```

To further examine the q-vectors that are suggested to be modified, you can draw mesa plots:

```{r,eval=FALSE}
plot(Qv, item = 9)
```

4. Perform item-level model selection to see if the saturated G-DINA model can be simplified:
```{r,eval=FALSE}
mc <- modelcomp(est)
mc
```

5. Assess model-data fit at test and item levels:
```{r,eval=FALSE}
# test level absolute fit
mft <- modelfit(est)
mft
# item level absolute fit
ift <- itemfit(est)
ift
summary(ift)
plot(ift)
```

6. Calculate classification accuracy
```{r,eval=FALSE}
CA(est)
```

7. Regress estimated classifications on user covariates using the three-step ML correction.

The bundled real-data examples do not include respondent-level covariates, so in the code below the covariates are just placeholders for user-supplied background variables collected outside the response matrix. The GDINA fit is real, based on the ECPE data, while the covariate frame `Z` is only used to illustrate the `ThreeStepCov()` workflow.

```{r,eval=FALSE}
dat <- realdata_ECPE$dat
Q <- realdata_ECPE$Q
ecpe_fit <- GDINA(dat, Q, model = "GDINA")

set.seed(123)
N <- nrow(dat)
Z <- data.frame(
  age = round(rnorm(N, mean = 21, sd = 3)),
  female = rbinom(N, 1, 0.55),
  program = factor(sample(c("intensive", "regular", "self-study"), N, replace = TRUE))
)

# Attribute-level regression with the three-step correction
att_reg <- ThreeStepCov(
  ecpe_fit,
  ~ age + female + program,
  data = Z,
  level = "attribute"
)
att_reg$results[[1]]$corrected$table
att_reg$results[[1]]$naive$table

# Profile-level regression for the estimated latent classes
profile_reg <- ThreeStepCov(
  ecpe_fit,
  ~ age + female + program,
  data = Z,
  level = "profile"
)
head(profile_reg$results$corrected$table)
```

At the attribute level, each element of `att_reg$results` contains both the naive logistic regression and the corrected regression for one attribute. At the profile level, `profile_reg$results$corrected$table` returns the multinomial logit coefficients relative to the reference latent profile.

## More Examples

* [G-DINA model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/GDINA_example.html)
* [Model estimation and diagnostics](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/CDMAnalysis_example.html) 
* [Sequential model estimation and diagnostics](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/sGDINA_analysis_example.html)
* [LCDM model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/LCDM_example.html) 
* [Higher-order GDINA model estimation](https://wenchao-ma.github.io/GDINA/articles/OnlineExercises/HOGDINA_example.html)  

If you would like to contribute an example to this website, please send me your .Rmd file.
